---
title: "Systematic Literature Review and Meta-analyses of Impact of Water, Sanitation, and Hygiene on the Transmission of Choler and Typhoid Fever."
output: html_notebook
editor_options:
chunk_output_type: console
---

```{r}
# load libraries
library (brms)
library (dplyr)
library (forcats)
library (ggplot2)
library (ggridges)
library (glue)
library (metafor)
library (readxl)
library (rstan)
library (stringr)
library (tidybayes)
library (googlesheets4)
library (data.table)

# remove all objects from workspace
rm (list = ls ())
setwd("~/GitHub/WASH_typhoid")

# start time
start_time <- Sys.time ()
print (paste0 ("start time = ", start_time))
```



### Importing data from Google Sheets
```{r}

# import data -- combined data
# input_articles <- read_excel("data/input_articles.xlsx")
dat <- read_sheet('https://docs.google.com/spreadsheets/d/1ArKNfpa124oPdrbGK1-9PnzRWWvffsxgOUfgCVs_p2Q/edit#gid=0', col_types = "ccccccccccccccc")

1

```



### Parsing texts to extract upper and lower bounds of the confidence interval
```{r}
# parsing test
ci_text <- "2.5-21"
ci_text <- ".19-1.0"
lb <- grep("[0-9]*", ci_text, value = T)
# as.numeric(str_extract_all(ci_text,"\\(?[0-9,.]+\\)?")[[1]])
as.numeric(str_extract_all(ci_text, "[0-9.]+")[[1]])
```



### Standard error computation
```{r}
calc_se <- function(lower, upper){
  se <- (upper - lower) / (2 * qnorm(0.975))
  return (se)
}
```



### Data cleaning
```{r}

# rename the columns
dat <- dat %>% rename("mOR"  = "Crude OR",
                      "mCI"  = "Crude OR CI",
                      "aOR"  = "Adjusted OR",
                      "aCI"  = "Adjusted OR CI")

dat <- data.table(dat)

# remove the blank rows
dat <- dat[!(is.na(dat$mOR) & is.na(dat$aOR)),]

# remove values not classified
dat <- dat[!(dat$`JMP WASH Category` == "Not Classified")]

# remove reference values
dat <- dat[(dat$mOR != "ref" | is.na(dat$mOR)) & (dat$aOR != "ref" | is.na(dat$aOR)),]

# remove studies not using blood culture to confirm typhoid
# check the number of studies not using blood culture to confirm typhoid 
# dat %>% filter(dat$Culture == "No") %>% count(Author, Culture)
dat <- dat[dat$Culture != "No",]

```



### Create analysis baseline
```{r}

# inspect JMP WASH Category
unique(dat$`JMP WASH Category`)

dat %>% count(`JMP WASH Category`)

category_meta <- c("Water Source (Improved)",
                   "Water Source (Surface water)",
                   "Water Treatment (Treated water)",
                   "Water Treatment (Untreated water)",
                   "Sanitation (Open defecation)",
                   "Hygiene (Basic)", 
                   "Hygiene (Limited)")

# create baseline data
OR <- as.numeric(ifelse(is.na(dat$aOR), dat$mOR, dat$aOR))
ci_text <- ifelse(is.na(dat$aCI), dat$mCI, dat$aCI)
ci <- str_extract_all(ci_text, "[0-9.]+")
lower <- as.numeric(sapply(ci, function(x) x[1]))
upper <- as.numeric(sapply(ci, function(x) x[2]))
se <- calc_se(lower = log(lower), upper = log(upper))
input_data <- data.frame(category = dat$`JMP WASH Category`, 
                         study    = dat$Study, 
                         log_OR   = log(OR), 
                         SE       = se)
```



### Frequentist Meta-Analysis
```{r}

meta_freq <- function(input_category, image_png){

input_dat_freq <- input_data %>% filter(input_data$category == input_category)
# input_dat_freq <- input_data %>% filter(input_data$category == "Water Source (Improved)")
# input_dat_freq <- input_data %>% filter(input_data$category == "Water Treatment (Untreated water)")

# frequentist random-effect meta-analysis
res <- rma(yi = log_OR, sei = SE, data = input_dat_freq)

summary <- summary(res)

# forest plot
# forest(res, atransf=exp)

# save funnel plot
png(file = file.path("figures", image_png), width = 1600, height = 1400, res = 300)

funnel (res, xlab = "Log odds ratio") # Log odds ratio with standard error of log odds ratio
# funnel (res, atransf = exp, xlab = "Odds ratio")
title(input_category)

dev.off()

return(summary)}

Water_Source_Improve_fre <- 
  meta_freq(input_category = "Water Source (Improved)",
            image_png      = "Water Source (Improved).png")

Water_Source_Surface_water_fre <- 
  meta_freq(input_category = "Water Source (Surface water)",
            image_png      = "Water Source (Surface water).png")

Water_Treatment_Treated_water_fre <-
  meta_freq(input_category = "Water Treatment (Treated water)",
            image_png      = "Water Treatment (Treated water).png")

Water_Treatment_Untreated_water_fre <-
  meta_freq(input_category = "Water Treatment (Untreated water)",
            image_png      = "Water Treatment (Untreated water).png")
  
Sanitation_Open_defecation_fre <-
  meta_freq(input_category = "Sanitation (Open defecation)",
            image_png      = "Sanitation (Open defecation).png")

Hygiene_Basic_fre <-
  meta_freq(input_category = "Hygiene (Basic)",
            image_png      = "Hygiene (Basic).png")

Hygiene_Limited_fre <-
  meta_freq(input_category = "Hygiene (Limited)",
            image_png      = "Hygiene (Limited).png")

```



### Bayesian Meta-Analysis
```{r}
meta_baye <- function(input_category){

input_dat_baye <- input_data %>% filter(input_data$category == input_category)

priors <- c(prior(normal(0,1), class = Intercept),
            prior(cauchy(0,0.5), class = sd))

m_brm <- brm(log_OR|se(SE) ~ 1 + (1|study),
             data = input_dat_baye,
             prior = priors,
             iter = 10000)

return(m_brm)}

```



### Prepare the data to generate the plot 
```{r}
pooled_effect_draws <- function(m_brm){
# extract the posterior distribution for each study
study.draws <- spread_draws(m_brm, r_study[study,], b_Intercept) %>% 
  mutate(b_Intercept = r_study + b_Intercept)

# generate the distribution of the pooled effect
pooled.effect.draws <- spread_draws(m_brm, b_Intercept) %>% 
  mutate(study = "Pooled Effect")

forest.data <- bind_rows(study.draws, 
                         pooled.effect.draws) %>% 
   ungroup() %>%
  # clean the study labels
   mutate(study = str_replace_all(study, ".et.al.", " et al.")) %>% 
  # reorder the study factor levels by effect size (high to low).
   mutate(study = reorder(study, b_Intercept)) %>% 
  # use exp() to revert it to Odd Ratio
  mutate(b_Intercept = exp(b_Intercept))

return(forest.data)}
```



### Function to generate forest plot
```{r}
create_forestplot <- function(m_brm,
                              forest.data,
                              x_limit,
                              title_name){

  # display the effect size
  forest.data.summary <- group_by(forest.data, study) %>% 
  median_qi(b_Intercept)
  
  forestplot <- ggplot(aes(b_Intercept, relevel(study, "Pooled Effect", after = Inf)),
       data = forest.data) +
  
  # Add vertical lines for pooled effect and CI
  geom_vline(xintercept = data.table(forest.data.summary)[study == "Pooled Effect", b_Intercept], 
             color = "grey", size = 0.6) +
  geom_vline(xintercept = data.table(forest.data.summary)[study == "Pooled Effect", c(.lower, .upper)],
             color = "grey", linetype = 2) +
  geom_vline(xintercept = 1, color = "black", size = 0.6) +
  scale_x_continuous(limits = c(0, x_limit)) +
  
  # Add posterior distribution for each study individually and pooled effect
  stat_pointinterval(.width = c(.8, .95)) +
    
  # Add text and labels
  geom_text(data = mutate_if(forest.data.summary, 
                             is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), 
        x = Inf), hjust = "inward") +
  labs(x = "Odd Ratio [95% Credible Interval]",
       y = element_blank()) +
  theme_bw() +
  
  ggtitle(paste(title_name)) +
  theme(plot.title = element_text(hjust=0.5, vjust=2, size = 15))
  
  return(forestplot)
  } # end of function -- create_forest_plot

```



### Combine functions to generate and save forest plot
```{r}

forest_plot <- function(input_category,
                        x_limit){
  
m_brm_input <- meta_baye(input_category)

forest_data <- pooled_effect_draws(m_brm_input)

forestplot  <- create_forestplot(m_brm       = m_brm_input,
                                 forest.data = forest_data, 
                                 x_limit     = x_limit,
                                 title_name  = input_category)

 # save plot
  ggsave (filename = paste(input_category, ".png"),
          path = "figures",
          width = 7,
          height = 3.5,
          dpi = 600)

return(m_brm_input)}
```



### Grenerate forest plots
```{r}
# JMP WASH categories included in the meta-analysis
category_meta

dat %>% count(`JMP WASH Category`)

# lapply(category_meta, forest_plot)
# R session aborted

# generate forest plot for each category individually
set.seed(3)

Water_Source_Improve <- 
forest_plot(input_category = "Water Source (Improved)",
            x_limit        = 5)

Water_Source_Surface_water <-
forest_plot(input_category = "Water Source (Surface water)",
            x_limit        = 6)

Water_Treatment_Treated_water <-
forest_plot(input_category = "Water Treatment (Treated water)",
            x_limit        = 3)

Water_Treatment_Untreated_water <-
forest_plot(input_category = "Water Treatment (Untreated water)",
            x_limit        = 12)

Sanitation_Open_defecation <-
forest_plot(input_category = "Sanitation (Open defecation)",
            x_limit        = 6)

Hygiene_Basic <-
forest_plot(input_category = "Hygiene (Basic)",
            x_limit        = 3)

Hygiene_Limited <-
forest_plot(input_category = "Hygiene (Limited)",
            x_limit        = 10)
```



### Assessing convergence and model validity
```{r}
# R hat -- the Potential Scale Reduction Factor (PSRF) -- of the parameter estimates should be lower than 1.01
summary(Water_Source_Improve)
summary(Water_Source_Surface_water)
summary(Water_Treatment_Treated_water)
summary(Water_Treatment_Untreated_water)
summary(Sanitation_Open_defecation)
summary(Hygiene_Basic)
summary(Hygiene_Limited)

# posterior predictive checks
pp_check(Water_Source_Improve, ndraws = 100)
pp_check(Water_Source_Surface_water, ndraws = 100)
pp_check(Water_Treatment_Treated_water, ndraws = 100)
pp_check(Water_Treatment_Untreated_water, ndraws = 100)
pp_check(Sanitation_Open_defecation, ndraws = 100)
pp_check(Hygiene_Basic, ndraws = 100)
pp_check(Hygiene_Limited, ndraws = 100)

```



```{r}
# end time
end_time <- Sys.time ()
print (paste0 ("end time = ", end_time))
print (Sys.time () - start_time)

```
